Finiteness in Cubical Type Theory

### 2020-06-23 sections/finiteness-predicates.tex
added one countdown example

\begin{lemma} \label{split-enum-sigma}
  Split enumerability is closed under \(\Sigma\).
  \begin{equation}
    \frac{
      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \; \; \; \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U(x))
    }{
      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\Sigma(x : A) , U(x))
    }
  \end{equation}
\end{lemma}
\begin{proof}
  Let \(A\) be a type which is split enumerable, and \(U\) be a type family over
  \(A\) which is split enumerable at every point.
  Formally, we have the following proofs:
  \begin{align}
    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_A &: \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \\
    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_U &: \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U(x))
  \end{align}

@@

\begin{agdalisting}
  This pattern of applying a function to each element in a list and
  concatenating the result is of course well-known in functional programming,
  and is in fact the pattern that makes lists a monad.
  While this insight isn't strictly relevant to our work here, it does mean
  the implementation of this function can use Agda's do notation, resulting
  in the following extremely clean implementation:
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sup-sigma}
\end{agdalisting}

>>>>

\begin{lemma} \label{split-enum-sigma}
  Split enumerability is closed under \(\Sigma\).
  \begin{equation}
    \frac{
      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\;A \; \; \; (x : A) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U\;x)
    }{
      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{\ensuremath{\Sigma [}}\;x \AgdaDatatype{:} A \AgdaDatatype{]} U\;x)
    }
  \end{equation}
\end{lemma}
\begin{proof}
  Let \(A\) be a type which is split enumerable, and \(U\) be a type family over
  \(A\) which is split enumerable at every point.
  Formally, we have the following proofs:
  \begin{align}
    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_A &: \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \\
    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_U &: \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U(x))
  \end{align}

@@

This pattern of applying a function to each element in a list and
concatenating the result is of course well-known in functional programming,
and is in fact the pattern that makes lists a monad.
While this insight isn't strictly relevant to our work here, it does mean
the implementation of this function can use Agda's do notation, resulting
in the following extremely clean implementation:
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sup-sigma}
\end{agdalisting}

We now have two components we'll need for the proof that the countdown
transformation is finite.
The component we'll look at is step~\ref{countdown-operators}: selection of the
operators.
We'll first need a type representing the operators available to us.
\begin{agdalisting}
  \ExecuteMetaData[agda/Countdown.tex]{ops-def}
\end{agdalisting}
Proving that this type is finite takes much the same form as the proof of
finiteness for bool.

Next, we will need to build a proof of finiteness for vectors of length \(n\).
This uses the proof of finiteness for \(\Sigma\).
\begin{agdalisting}
  \ExecuteMetaData[agda/Countdown.tex]{vec-fin}
\end{agdalisting}
